---
title: "Attitude of Youth towards E-cigarettes"
author: Shengjie Sun
output: html_notebook
---

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(GGally)
setwd("C:/Users/sheng/OneDrive/CU Second Semester/Statistical Graphics/Final Project")

questions_num <- readr::read_csv("./output/attitude/Number of Questions.csv")
questions_num$Type <- factor(questions_num$Type, levels = c("Indirectly Related", "Directly Related"))

NYTS_2018_full <- readr::read_csv("./data/2018-NYTS/nyts2018.csv") %>% select(matches("Q(1|2|3|5)[A-Z]?$"),  contains('Q33'), 'Q27', 'Q35', 'Q68', 'Q69', 'Q28','Q29','Q30','Q31','Q34','Q36','Q37') %>%
  mutate(Q33A = ifelse(Q33A==1, "A", Q33A)) %>% mutate(Q33B = ifelse(Q33B==1, "B", Q33B)) %>%
  mutate(Q33C = ifelse(Q33C==1, "C", Q33C)) %>% mutate(Q33D = ifelse(Q33D==1, "D", Q33D)) %>%
  mutate(Q33E = ifelse(Q33E==1, "E", Q33E)) %>% mutate(Q33F = ifelse(Q33F==1, "F", Q33F)) %>%
  mutate(Q33G = ifelse(Q33G==1, "G", Q33G)) %>% mutate(Q33H = ifelse(Q33H==1, "H", Q33H)) %>%
  mutate(Q33I = ifelse(Q33I==1, "I", Q33I)) %>% mutate(Q33J = ifelse(Q33J==1, "J", Q33J)) %>%
  #mutate(Q33Fake = coalesce(Q33A,Q33B,Q33C,Q33D,Q33E,Q33F,Q33G,Q33H,Q33I,Q33J)) %>%
  mutate(Q33 = stringr::str_replace_all(paste(Q33A,Q33B,Q33C,Q33D,Q33E,Q33F,Q33G,Q33H,Q33I,Q33J, sep = ""),"NA","")) %>%
  mutate(Q5A = ifelse(Q5A==1, "A", Q5A)) %>% mutate(Q5B = ifelse(Q5B==1, "B", Q5B)) %>%
  mutate(Q5C = ifelse(Q5C==1, "C", Q5C)) %>% mutate(Q5D = ifelse(Q5D==1, "D", Q5D)) %>%
  mutate(Q5E = ifelse(Q5E==1, "E", Q5E)) %>% 
  select(matches("Q(1|2|3)$"), 'Q27','Q28','Q29','Q30','Q31','Q34','Q35','Q36','Q37','Q68','Q69', contains('Q33')) %>%
  filter(Q2!='*') %>% filter(Q3!='*') %>% filter(Q34!='*') %>% filter(Q27!='*') %>% 
  filter(Q35!='*') %>% filter(Q68!='*') %>% filter(Q69!='*') %>% filter(Q37!='*') %>% 
  filter(Q29!='**') %>% filter(Q1!='**')
```

## Number of Questions
```{r}
ggplot(data = questions_num) + 
  geom_col(aes(x=Year, y=Number, fill=Type)) + 
  scale_fill_manual(values = c("gray75", "gray45"))
```

## Curiousity and trying e-cigarette
Q3. What grade are you in?
Q27. Have you ever been curious about using an e-cigarette?
Q35. Do you think that you will try an e-cigarette soon?
Q36. Do you think you will use an e-cigarette in the next year?
Q37. If one of your best friends were to offer you an e-cigarette, would you use it?
```{r}
NYTS_2018_full %>% 
  filter(Q28 == 2) %>%
  select('Q27','Q3', 'Q35','Q36','Q37') %>% 
  mutate(Type = ifelse(Q27<=2, "Curious","Not Curious")) %>%
  mutate(Q3 = readr::parse_integer(Q3)-1.5) %>%
  sample_n(2000) %>%
  ggparcoord(columns = 2:5, scale = "globalminmax", splineFactor = 10, alphaLines = 0.5, groupColumn = 6) 
```

## 33. What are the reasons you have used ecigarettes? (Select one or more)
A. I have never tried an e-cigarette
### B. Friend or family member used them
C. To try to quit using other tobacco products, such as cigarettes
D. They cost less than other tobacco products, such as cigarettes
E. They are easier to get than other tobacco products, such as cigarettes
F. Famous people on TV or in movies use them
### G. They are less harmful than other forms of tobacco, such as cigarettes
### H. They are available in flavors, such as mint, candy, fruit, or chocolate
I. They can be used in areas where other tobacco products, such as cigarettes, are not allowed
### J. I used them for some other reason
```{r}
Reasons <- NYTS_2018_full %>% filter(is.na(Q33A)) %>% select(matches("Q(1|2|3)$"), contains('Q33'))
Reasons %>% count(Q33) %>% filter(n>=10) %>% arrange(desc(n)) %>% filter(n!=75)
```

